Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-29 paper.tex
more edits

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu out of the fuel fab and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The colored curves show Pu inventory
available for making fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases 2 and 4, the individual reactor scenarios (cases 1 and 3) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

Quarterly time steps have less frequent but larger impulse material flows.
These larger flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level during
the shortage.  This can be clearly seen in the case 4 plot which has a higher
standing inventory than case 2 during the fuel shortage years.  The case 4 Pu
inventory is unable to drop below approximately 100 tonnes, which indicates
out-flows to be roughly equal to the Pu flow into inventory each time step
during those years.  This can be confirmed by the in-flow and inventory curves
touching at those points.  Case 2 Pu inventory is similarly limited at about
the 30 tonne level.  As expected, the case 4 withdrawals are also three times
higher case 2 withdrawals exactly matching the time step duration ratio
between the two cases.

@@

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Counter-intuitively, lower separated plutonium
inventory during the shortage generally indicates fewer unfueled reactors.
Lower inventory levels mean available Pu is being utilized more efficiently
rather than sitting idle.  Cases 3 and 4 with their longer time steps show
higher inventories indicating a more severe shortage.  Post shortage, they
also fall a bit behind on building up Pu stocks due to the delayed
contributions of more unfueled reactors to the separated Pu pool.  The
reactors in these scenarios request fresh fuel only when they need it. While
this is an unrealistic behavior for real-world facilities, it is closer to a
globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
would also have the same effect of increasing idling plutonium quantities and
making the fuel shortage worse overall.

@@

Also notable is that different modeling choices (e.g. time step duration,
facility discretization, etc.) may have varying levels of significance
depending on the scenario/context they are operating in.  As shown in Figure
\ref{fig:puinv-compare} Before the fuel shortage begins, differences between
each of the four cases are relatively minor.  However, after the shortage
begins around year 100, separated Pu inventories vary much more significantly
between the different cases.

@@

Different modeling choices can have a significant impact on the outcome of a
simulation.  With discrete reactor modeling many factors must be considered in
order to ensure the integrity of conclusions.  Individual reactor outage
modeling can be very useful for certain analyses, but it can reduce result
quality if things such as cycle staggering are handled appropriately.  The
fleet and discreet modeling done here is itself multi-dimensional.  If
individual outages are not of interest, reactor power could be lowered and the
outage reduced to zero duration making the capacity factor part of the
reactor's power capacity.  The cycle staggering effects present when modeling
reactors individually can provide valuable insight not possible with the
fleet-based reactor modeling.  Starting reactors following shortage induced
outages should not follow the natural pacing of fuel availability, otherwise
refueling cycles become unsatisfactorily synchronized.

Increasing the time steps increases facilities' standing inventory requirement
for avoiding supply constraints.  Facilities that have unconstrained supply
and very large inventory buffers, are not affected by the time step in this
way.  Facilities can provide no more inventory than they can hold on a single
time step.  So increasing time step duration for a facility with unconstrained
supply generally means higher standing inventories are required in order to
achieve equal throughput. Even when facilities do not have individual
inventory or throughput limits, the time step can still have a significant
effect. In recycle loops there is a constraint on how fast material is
introduced through the loop (i.e. the discharge rate of spent fuel from
reactors).  The \emph{inventory drawdown effect} can reduce the aggregate
throughput of the recycling loop.

The impact of modeling assumptions depends heavily on the metrics of interest.
For example, someone interested in fuel-shortage driven reactor outages could
observe a factor of four span among the four cases shown in the detail
comparison.  However, someone more interested in the overall transition speed
would not see as significant a discrepancy.

In an unconstrained supply setting, varying the time step duration or
discretization of facility modeling have a relatively minor impact on the
overall simulation results.  Differences become more significant in material
constrained regimes.  Robustness in constrained regimes is especially useful
in the context of sensitivity and optimization analysis.

>>>>

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu out of the fuel fab and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The other colored curves show Pu inventory
available for making fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases 2 and 4, the individual reactor scenarios (cases 1 and 3) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

Quarterly time steps have less frequent but larger impulse material flows.
The larger in-flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level during
the shortage.  This can be clearly seen in the case 4 plot which has a higher
standing inventory than case 2 during the fuel shortage years.  The case 4 Pu
inventory is unable to drop below approximately 100 tonnes, which indicates
out-flows to be roughly equal to the Pu flow into inventory each time step
during those years.  This can be confirmed by the in-flow and inventory curves
touching at those points.  Case 2 Pu inventory is similarly limited at about
the 30 tonne level.  As expected, the case 4 withdrawals are also three times
higher case 2 withdrawals exactly matching the time step duration ratio
between the two cases.

@@

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
inventory during the shortage generally indicates fewer unfueled reactors.
Lower inventory levels mean available Pu is being utilized more efficiently
rather than sitting idle.  Cases 3 and 4 with their longer time steps show
higher inventories indicating a more severe shortage.  Post shortage, they
also fall a bit behind on building up Pu stocks due to the delayed
contributions of more unfueled reactors to the separated Pu pool.  The
reactors in all the scenarios request fresh fuel only when they need it. While
this is an unrealistic behavior for real-world facilities, it is closer to a
globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
would also have the same effect of increasing idling plutonium quantities and
would make the fuel shortage worse overall.

@@

Also notable is that different modeling choices (e.g. time step duration,
facility discretization, etc.) may have varying levels of significance
depending on the scenario/context they are operating in.  As shown in Figure
\ref{fig:puinv-compare} before the fuel shortage begins, differences between
each of the four cases are relatively minor.  However, after the shortage
begins near year 100, separated Pu inventories vary much more significantly
between the different cases.

@@

Different modeling choices can have a significant impact on the outcome of a
simulation.  With discrete reactor modeling many factors must be considered in
order to ensure the integrity of conclusions.  Individual reactor outage
modeling can be very useful for certain analyses, but it can reduce result
quality if things such as cycle staggering are not handled appropriately.  The
fleet and discreet modeling done here is itself multi-dimensional. If fuel
sharing is of interest, but cycle staggering is not, individual reactors can
be used with the power capacity lowered and the outage reduced to zero
duration making the capacity factor included in the reactor's power capacity.
Modeling reactors individually can provide valuable insight not possible with
the fleet-based reactor modeling.  For example, bringing reactors back online
following shortage-induced outages should not follow the natural pacing of
fuel availability, otherwise refueling cycles become unsatisfactorily
synchronized.

Increasing the time steps increases facilities' standing inventory requirement
for avoiding supply constraints.  Facilities can provide no more inventory
than they can hold on a single time step.  So increasing time step duration
for a facility with unconstrained supply generally means higher standing
inventories are required in order to achieve equal throughput. Facilities that
have unconstrained supply and very large inventory buffers, are not affected
meaningfully by the time step. Even when facilities do not have individual
inventory or throughput limits, the time step can still have a significant
effect. In recycle loops there is a constraint on how fast material is
introduced through the loop (i.e. the discharge rate of spent fuel from
reactors).  The \emph{inventory drawdown effect} can impact the aggregate
throughput of the recycling loop.

The impact of modeling assumptions depends heavily on the metrics of interest.
For example, someone interested in fuel-shortage driven reactor outages would
observe a factor of four span among the four cases shown in the detail
comparison.  However, someone more interested in the overall transition
mechanics would not see as significant a discrepancy.  In an unconstrained
supply setting, varying the time step duration or discretization of facility
modeling have a relatively minor impact on the overall simulation results.
Differences become more significant in material constrained regimes.
Robustness in constrained regimes is especially useful in the context of
sensitivity and optimization analysis.
